The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-05-30 paper/ms.tex
down to integrated line intensity

The central results of this paper are summarized in Figures
\ref{fig:CentralSpec} and \ref{fig:HomSpec}. They show 2D histograms
of the escape frequency $x$ and outgoing angle $\theta$ parameterized by
$|\mu|$. Constructing the emission line with photons around a value
of $\mu$ would gives us the emission detected by an observer located at an
angle $\theta$ with respect to the rotation axis. We have veryfied
that the solutions are indeed symmetric with respect to $\mu=0$.

@@

If the viewing angle is aligned witht the rotation axis, $|\mu|\sim
1$, the \ly line keeps a double peak with some changes as a function
of the rotational velocity that are easier to spot in models where the
\ly photons are homogeneously distributen than in the case of central
emission.

However, for a line of sight perpendicular to the rotation
axis, $|\mu|\sim 0$, the impact of rotation is larger.  This is clear
in  Figure \ref{fig:differentvelocities} where we present the
different line morphologies for $|\mu|\sim 0$ for both the homogeneous
and central configurations. The panels have the same distribution as
Figures \ref{fig:CentralSpec} and \ref{fig:HomSpec}.

For this case of $|\mu|\sim 0$, there are three clear effects on the
line's morphology as the rotational velocity increases. First, the
line broadens; second, the double peaks reduce their intensity; and
third, the intensity at the line center rises. The last two effects
merge to give the impression that the double peaks are merged into one
at high rotational velocities, a result that is evident for the
homogeneously distributed sources as shown in the dashed lines of
Figure \ref{fig:differentvelocities}.

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f5.pdf}
\end{center}
\caption{Integrated flux distribution as a function of the
  viewing angle as parameterized by $\mu$. We show the results for the
  homogeneous source distribution in models with rotational velocities
  $0$\kms, $300$\kms and  optical depths $10^5$, $10^6$ and $10^7$.
  The distributions are flat, meaning that the integrated flux for all
  viewing angles is the same. The results for the central sources are
  similar. 
\label{fig:muhisto}}
\end{figure}

We consider possible variations in the integrated flux with respect to
$\theta$. We define the normalized flux seen by an observer at an
angle $\mu$ by: \begin{equation}
F(\mu) = \frac{2\Delta N}{N\Delta \mu}, 
\end{equation} 
%
where $\mu=\cos\Theta$, $N$ is the total number of outgoing photons,
$\Delta N$ is the number of photons in an angular bin $\Delta
\Theta$. This definition satisfies the condition
$\int_{-1}^{1}F(\mu)d\mu/2=1$.

Figure \ref{fig:muhisto} shows a flat $F(\mu)$ for all models. It
means that in spite of the different morphologies, the integrated flux
is the same for all sight angles.

@@

{\bf Galaxy simulations with gas rotation}. \cite{Verhamme12} studied \ly
line emission in two high resolution simulations of individual
galaxies. The main purpose of their study was to asses the impact of
two different Inter-Stellar Medium (ISM) prescriptions. However, each
simulated galaxy had a disc structure with a clear rotation pattern in
the ISM and inflowing gas from the circum-galactic region. The configuration
had an axial symmetry and they reported a strong dependence of both
the escape fraction and the total line intensity as a function of the
$\theta$ angle. In our results, none of these two quantities has a
dependence either on the inclination angle or the rotational
velocity. We suggest that he effect reported by \cite{Verhamme12} is
consistent with being a consequence of the different hydrogen optical
depth for different viewing angles and not as an effect of the bulk
rotation.

@@

The definitive and clear impact of rotation on the \ly morphology
suggests that this is an effect that should be taken into accunt at
the moment of interpreting high resolution spectroscopic data. To this
end, we defer a discussion of observed lines features in the context of gas
bulk rotation to a future publication (Remolina-Gutierrez et al., in
preparation).

>>>>

The central result of this paper are summarized in Figures
\ref{fig:CentralSpec} and \ref{fig:HomSpec}. They show 2D histograms
of the escape frequency $x$ and outgoing angle $\theta$ parameterized by
$|\mu|$. Constructing the emission line with photons around a value
of $\mu$ would gives us the emission detected by an observer located at an
angle $\theta$ with respect to the rotation axis. We have veryfied
that the solutions are indeed symmetric with respect to $\mu=0$.

@@

If the viewing angle is aligned witht the rotation axis, $|\mu|\sim
1$, the \ly line keeps a double peak with minor changes with
increasing rotational velocity. However, for a line of sight
perpendicular to the rotation axis, $|\mu|\sim 0$, the impact of
rotation is larger. This is clear in  Figure
\ref{fig:differentvelocities} where we present the different line
morphologies for $|\mu|\sim 0$ for both the homogeneous and central
configurations. The panels have the same distribution as Figures
\ref{fig:CentralSpec} and \ref{fig:HomSpec}.

In the case of $|\mu|\sim 0$ there are three clear effects on the
line morphology as the rotational velocity increases. First, the
line broadens; second, the double peaks reduce their intensity; and
third, the intensity at the line center rises. The last two effects
merge to give the impression that the double peaks are merged into one
at high rotational velocities, a result that is evident for the
homogeneously distributed sources as shown in the dashed lines of
Figure \ref{fig:differentvelocities}.

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.4\textwidth]{f5.pdf}
\end{center}
\caption{Integrated flux distribution as a function of the
  viewing angle as parameterized by $\mu$. Continous (dashed)
  correspond to central (homogeneous) source distribution.
  The models correspond to an optical depth of $\tau_{\rm
    H}=10^5$ and rotational velocities of $100$\kms, $200$\kms and
  $300$\kms. The distributions are flat, meaning that the integrated
  flux for all viewing angles is the same.
\label{fig:muhisto}}
\end{figure}

Besides the change in morphology, we now consider possible variations
in the integrated flux with respect to the viewing angle $\theta$. We
define the normalized flux seen by an observer at an angle $\mu$ by:

\begin{equation}
F(\mu) = \frac{2\Delta N}{N\Delta \mu}, 
\end{equation} 
%
where $\mu=\cos\Theta$, $N$ is the total number of outgoing photons,
$\Delta N$ is the number of photons in an angular bin $\Delta
\Theta$. This definition satisfies the condition
$\int_{-1}^{1}F(\mu)d\mu/2=1$.  In the case of perfect spherical
simmetry one expects a flat distribution of $F(\mu)=1$.

Figure \ref{fig:muhisto} shows the results for a selection of models
with $\tau_{\rm H}=10^{5}$, different rotational velocities and both
for the central and homogeneous source distributions. The
distributions do not show any trend, beyond some fluctuations below
the $2\%$ level. This result implies that the integrated flux is
constant for all viewing angles and all the possible models.

@@

{\bf Galaxy simulations with gas rotation}. \cite{Verhamme12} studied \ly
line emission in two high resolution simulations of individual
galaxies. The main purpose of their study was to assess the impact of
two different Inter-Stellar Medium (ISM) prescriptions. However, each
simulated galaxy had a disc structure with a clear rotation pattern in
the ISM and inflowing gas from the circum-galactic region. The configuration
had an axial symmetry and they reported a strong dependence of both
the escape fraction and the total line intensity as a function of the
$\theta$ angle. In our results, none of these two quantities has a
dependence either on the inclination angle or the rotational
velocity. We suggest that he effect reported by \cite{Verhamme12} is
consistent with being a consequence of the different hydrogen optical
depth for different viewing angles and not as an effect of the bulk
rotation.

@@

The definitive and clear impact of rotation on the \ly morphology
suggests that this is an effect that should be taken into accunt at
the moment of interpreting high resolution spectroscopic data. To this
end, we defer a discussion of observed lines features in the context of gas
bulk rotation to a future publication (Remolina-Gutierrez et al., in
prep.).
